% This LaTeX document was generated using the LaTeX backend of PlDoc,
% The SWI-Prolog documentation system

% This file was generated from swipl-devel/man with:
% $ swipl -q -f libtotex.pl -g libtotex -- 'library(clpfd)'

\begin{summarylist}{ll}
\oppredsummary{\Scand}{2}{yfx}{720}{P and Q hold.}
\oppredsummary{\Sclt}{2}{xfx}{700}{The arithmetic expression X is less than Y.}
\oppredsummary{\Slimplies}{2}{yfx}{750}{Q implies P.}
\oppredsummary{\Scequal}{2}{yfx}{760}{P and Q are equivalent.}
\oppredsummary{\Sceq}{2}{xfx}{700}{The arithmetic expression X equals Y.}
\oppredsummary{\Scle}{2}{xfx}{700}{The arithmetic expression X is less than or equal to Y.}
\oppredsummary{\Srimplies}{2}{xfy}{750}{P implies Q.}
\oppredsummary{\Scgt}{2}{xfx}{700}{Same as Y \Sclt{} X.}
\oppredsummary{\Scge}{2}{xfx}{700}{Same as Y \Scle{} X.}
\oppredsummary{\Scnot}{1}{fy}{710}{Q does _not_ hold.}
\oppredsummary{\Scnot}{2}{fy}{710}{Either P holds or Q holds, but not both.}
\oppredsummary{\Scor}{2}{yfx}{740}{P or Q holds.}
\oppredsummary{\Scne}{2}{xfx}{700}{The arithmetic expressions X and Y evaluate to distinct integers.}
\predicatesummary{all_different}{1}{Like all_distinct/1, but with weaker propagation.}
\predicatesummary{all_distinct}{1}{True iff Vars are pairwise distinct.}
\predicatesummary{automaton}{3}{Describes a list of finite domain variables with a finite automaton.}
\predicatesummary{automaton}{8}{Describes a list of finite domain variables with a finite automaton.}
\predicatesummary{chain}{2}{Zs form a chain with respect to Relation.}
\predicatesummary{circuit}{1}{True iff the list Vs of finite domain variables induces a Hamiltonian circuit.}
\predicatesummary{cumulative}{1}{Equivalent to cumulative(Tasks, [limit(1)]).}
\predicatesummary{cumulative}{2}{Schedule with a limited resource.}
\predicatesummary{disjoint2}{1}{True iff Rectangles are not overlapping.}
\predicatesummary{element}{3}{The N-th element of the list of finite domain variables Vs is V.}
\predicatesummary{fd_dom}{2}{Dom is the current domain (see in/2) of Var.}
\predicatesummary{fd_inf}{2}{Inf is the infimum of the current domain of Var.}
\predicatesummary{fd_size}{2}{Reflect the current size of a domain.}
\predicatesummary{fd_sup}{2}{Sup is the supremum of the current domain of Var.}
\predicatesummary{fd_var}{1}{True iff Var is a CLP(FD) variable.}
\predicatesummary{global_cardinality}{2}{Global Cardinality constraint.}
\predicatesummary{global_cardinality}{3}{Global Cardinality constraint.}
\oppredsummary{in}{2}{xfx}{700}{Var is an element of Domain.}
\predicatesummary{indomain}{1}{Bind Var to all feasible values of its domain on backtracking.}
\oppredsummary{ins}{2}{xfx}{700}{The variables in the list Vars are elements of Domain.}
\predicatesummary{label}{1}{Equivalent to labeling([], Vars).}
\predicatesummary{labeling}{2}{Assign a value to each variable in Vars.}
\predicatesummary{lex_chain}{1}{Lists are lexicographically non-decreasing.}
\predicatesummary{scalar_product}{4}{True iff the scalar product of Cs and Vs is in relation Rel to Expr.}
\predicatesummary{serialized}{2}{Describes a set of non-overlapping tasks.}
\predicatesummary{sum}{3}{The sum of elements of the list Vars is in relation Rel to Expr.}
\predicatesummary{transpose}{2}{Transpose a list of lists of the same length.}
\predicatesummary{tuples_in}{2}{True iff all Tuples are elements of Relation.}
\predicatesummary{zcompare}{3}{Analogous to compare/3, with finite domain variables A and B.}
\end{summarylist}

